### 2022-04-14 paper/supplement.tex

\documentclass{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{authblk}
\usepackage{color}
\usepackage{graphicx}
\usepackage{lineno}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{url}

\linenumbers

\title{Supplement to Detection of slow slip events using wavelet analysis of GNSS recordings}
\author[1]{Ariane Ducellier}
\author[2]{Kenneth C. Creager}
\author[2]{David A. Schmidt}
\affil[1]{Corresponding author. University of Washington, Department of Earth and Space Sciences, Box 351310, 4000 15th Avenue NE Seattle, WA 98195-1310}
\affil[2]{University of Washington, Department of Earth and Space Sciences}
\date{}

\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}

\begin{document}

\maketitle

\begin{figure}
\noindent\includegraphics[width=4.5cm, trim={0cm 0cm 0cm 0cm},clip]{figures/DS_10.eps}
\caption{Top: \textbf{Original data from GPS station PGC5 (black) and same data where displacement values have been artificially removed at two different times (2010.97 on the left and 2011.48 on the right) and replaced by the sum of a straight line and a Gaussian noise component (red)}. The corresponding ten details and smooths of the wavelet composition are shown in increasing levels for the original data (black) and for the \textbf{data with displacement values removed and replaced} by linear interpolation plus Gaussian noise (red).}
\label{pngfiguresample}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_mode_details_6-7-8.eps}
\caption{Same as top panel of Figure 9 from the main text, but using the GNSS time series where common modes have been removed before applying the wavelet transform.}
\label{pngfiguresample}
\end{figure}

\end{document}